Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-03 paper.tex
replaced CPS by ECPS
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time). They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances. Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs). Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).

The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, internet of things - IoT) in the latter. This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain. For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}. One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a CPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.

Thus, CPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems. As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

@@

Besides, when physical interaction with the real world is needed, which happens in CPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

@@

The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by CPSs, IoT, and ML. We present a brief discussion about models for CPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research. Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems and Section \ref{Newapp} addresses new applications, which take into account the new trends mentioned here. Finally, we conclude and describe future work in Section~\ref{conclusions}.

@@

Automated and formal verification have been employed to ensure the correctness and reliability of CPS during the last decade, as well as controller synthesis for CPS and HS \textcolor{red}{add citations}. A CPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

The overall formal and automated synthesis process for CPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the CPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the CPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent CPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible. On the other hand, that approach is not suitable for complex CPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.

CPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.

@@

Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for CPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for CPS.

@@

Given a class of CPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in CPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model CPS) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must be hold for CPS are related to safety, liveness, and reachability. In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to CPSs: security. Indeed, the essence of CPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

>>>>

Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real-time)~\cite{Kopetz11}. They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, control, and even replace every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances~\cite{xu15}. Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers.

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs). Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs). The presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain. For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}. One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.

Thus, ECPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems. As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

@@

Besides, when physical interaction with the real world is needed, which happens in ECPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, such models can be used to synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

@@

The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, counter-example guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research. Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of embedded \& cyber-physical systems and Section \ref{Newapp} addresses new applications, which take into account the new trends mentioned here. Finally, we conclude and describe future work in Section~\ref{conclusions}.

@@

Automated and formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}. A ECPS $S$ is a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

The overall formal and automated synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS, modeled according to the first step (\textit{modeling}), to meet the requirements and behave as desired, as confirmed by the verification step.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible. On the other hand, that approach is not suitable for complex ECPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.

ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.

@@

Rungger and Tabuada~\cite{Rungger16} and Tabuada et al.~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.

@@

Given a class of ECPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPS) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must be hold for ECPS are related to safety, liveness, and reachability. In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.
